A pipeline to analyse time-course gene expression data

### 2019-08-30 manuscript.tex
Deploy NelleV/2019timecourse-rnaseq-pipeline to github.com/NelleV/2019timecourse-rnaseq-pipeline.git:gh-pages

EAP:

\begin{itemize}
\tightlist
\item
  I got an error in installation script for packages:
\end{itemize}

\begin{verbatim}
epurdom-0:2019timecourse-rnaseq-pipeline epurdom$ Rscript scripts/install.R
Loading required package: devtools
Loading required package: usethis
Loading required package: ggfortify
Loading required package: ggplot2
Loading required package: kableExtra
Error in contrib.url(repos, "source") : 
  trying to use CRAN without setting a mirror
Calls: install.packages -> contrib.url
In addition: Warning message:
In library(package, lib.loc = lib.loc, character.only = TRUE, logical.return = TRUE,  :
  there is no package called <e2><80><98>kableExtra<e2><80><99>
Execution halted
\end{verbatim}

@@

We note that we could have time course data on either a single condition (to
identify genes changing over time) or
time course data on multiple conditions (such as the influenza dataset we are
considering), which will alter slightly the types of questions we are
interested in.
In what follows, we will focus on the situation
where we have multiple conditions. \textbf{EAP: Any reason one and not the other.
Maybe should change this sentence}

@@

\begin{tabular}{l}
\hline
Contrasts\\
\hline
M.0-VL.0\\
\hline
M.3-VL.3\\
\hline
M.6-VL.6\\
\hline
M.9-VL.9\\
\hline
M.12-VL.12\\
\hline
M.18-VL.18\\
\hline
M.24-VL.24\\
\hline
M.30-VL.30\\
\hline
M.36-VL.36\\
\hline
M.48-VL.48\\
\hline
M.60-VL.60\\
\hline
M.72-VL.72\\
\hline
M.120-VL.120\\
\hline
M.168-VL.168\\
\hline
\end{tabular}

\textbf{EAP: Not sure if \texttt{kable} is better than just printout?}

@@

\begin{tabular}{l|r|r|r|r|r|r}
\hline
  & M.0-VL.0\_pval & M.0-VL.0\_qval & M.0-VL.0\_lfc & M.3-VL.3\_pval & M.3-VL.3\_qval & M.3-VL.3\_lfc\\
\hline
NM\_009912 & 0.273 & 0.475 & 2.202 & 0.767 & 0.876 & 1.158\\
\hline
NM\_008725 & 0.851 & 0.924 & 2.028 & 0.139 & 0.302 & 1.457\\
\hline
NM\_007473 & 0.088 & 0.219 & 0.678 & 0.917 & 0.959 & 0.764\\
\hline
ENSMUST00000094955 & 0.008 & 0.037 & 1.186 & 0.935 & 0.969 & -0.633\\
\hline
NM\_001042489 & 0.419 & 0.621 & 1.212 & 0.761 & 0.872 & 0.599\\
\hline
NM\_008159 & 0.123 & 0.278 & 1.102 & 0.117 & 0.268 & 0.737\\
\hline
NM\_001013813 & 0.057 & 0.159 & 0.630 & 0.171 & 0.348 & 0.975\\
\hline
AK039774 & 0.015 & 0.059 & 0.637 & 0.062 & 0.171 & 0.971\\
\hline
NM\_013782 & 0.875 & 0.938 & 0.594 & 0.000 & 0.001 & 1.180\\
\hline
NM\_028622 & 0.767 & 0.876 & 1.648 & 0.946 & 0.974 & -1.865\\
\hline
\end{tabular}

@@

\begin{tabular}{l|r|r|r|r|r}
\hline
  & M-K:1 & M-K:7 & M-K:11 & M-K:14 & M-K:2\\
\hline
NM\_009912 & -0.0061424 & 0.0224948 & -0.0515947 & -0.1143973 & 0.0614619\\
\hline
NM\_008725 & 2.6547855 & 0.2734387 & -0.6882925 & 0.4718807 & -1.7463319\\
\hline
NM\_007473 & -0.1603623 & 0.1017655 & 0.5079573 & 0.0595773 & 0.4250185\\
\hline
ENSMUST00000094955 & 0.3505920 & 0.6463011 & 0.2564796 & 0.3599102 & 0.4177116\\
\hline
NM\_001042489 & 0.0889750 & 0.2321381 & -0.1891501 & 0.1786113 & 0.3320644\\
\hline
NM\_008159 & -0.3863770 & 0.0298395 & -0.0862755 & -0.3266351 & -0.1893062\\
\hline
\end{tabular}

@@

\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedleft}X>{\raggedleft}X}
\hline
  & Pathway & Percentage & Adj. p-value\\
\hline
04060 & Cytokine-cytokine receptor interaction & 34 & 0\\
\hline
04620 & Toll-like receptor signaling pathway & 40 & 0\\
\hline
04630 & Jak-STAT signaling pathway & 33 & 0\\
\hline
03050 & Proteasome & 53 & 0\\
\hline
04621 & NOD-like receptor signaling pathway & 47 & 0\\
\hline
04380 & Osteoclast differentiation & 34 & 0\\
\hline
04623 & Cytosolic DNA-sensing pathway & 45 & 0\\
\hline
04210 & Apoptosis & 37 & 0\\
\hline
04622 & RIG-I-like receptor signaling pathway & 39 & 0\\
\hline
05160 & Hepatitis C & 29 & 0\\
\hline
\end{tabu}

@@

\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedright}X>{\raggedleft}X}
\hline
  & GO ID & Description & Annotated & Significant & Expected & P-value & adj. p-value\\
\hline
5 & GO:0003002 & regionalization & 172 & 153 & 134.41 & 0.00026 & 0.2629\\
\hline
6 & GO:0048562 & embryonic organ morphogenesis & 152 & 135 & 118.78 & 0.00045 & 0.3323\\
\hline
7 & GO:0010970 & transport along microtubule & 82 & 75 & 64.08 & 0.00046 & 0.3323\\
\hline
8 & GO:0006805 & xenobiotic metabolic process & 49 & 47 & 38.29 & 0.00059 & 0.3729\\
\hline
9 & GO:0060173 & limb development & 92 & 84 & 71.89 & 0.00068 & 0.3820\\
\hline
10 & GO:0019395 & fatty acid oxidation & 53 & 51 & 41.42 & 0.00134 & 0.6194\\
\hline
11 & GO:0055114 & oxidation-reduction process & 536 & 461 & 418.84 & 0.00139 & 0.6194\\
\hline
12 & GO:0060070 & canonical Wnt signaling pathway & 149 & 131 & 116.43 & 0.00147 & 0.6194\\
\hline
13 & GO:0090596 & sensory organ morphogenesis & 134 & 118 & 104.71 & 0.00219 & 0.8517\\
\hline
14 & GO:0060411 & cardiac septum morphogenesis & 33 & 32 & 25.79 & 0.00295 & 1.0000\\
\hline
15 & GO:0010811 & positive regulation of cell-substrate ad... & 81 & 73 & 63.30 & 0.00384 & 1.0000\\
\hline
16 & GO:0042738 & exogenous drug catabolic process & 22 & 22 & 17.19 & 0.00437 & 1.0000\\
\hline
17 & GO:1901381 & positive regulation of potassium ion tra... & 22 & 22 & 17.19 & 0.00437 & 1.0000\\
\hline
18 & GO:0007368 & determination of left/right symmetry & 66 & 60 & 51.57 & 0.00536 & 1.0000\\
\hline
19 & GO:0060078 & regulation of postsynaptic membrane pote... & 66 & 60 & 51.57 & 0.00536 & 1.0000\\
\hline
\end{tabu}

@@

\begin{verbatim}
## R version 3.6.0 (2019-04-26)
## Platform: x86_64-pc-linux-gnu (64-bit)
## Running under: Ubuntu 19.04
## 
## Matrix products: default
## BLAS:   /usr/lib/x86_64-linux-gnu/blas/libblas.so.3.8.0
## LAPACK: /usr/lib/x86_64-linux-gnu/lapack/liblapack.so.3.8.0
## 
## locale:
## [1] C
## 
## attached base packages:
##  [1] stats4    parallel  splines   stats     graphics  grDevices utils    
##  [8] datasets  methods   base     
## 
## other attached packages:
##  [1] RColorBrewer_1.1-2   moanin_0.0.0         topGO_2.32.0        
##  [4] SparseM_1.77         GO.db_3.6.0          AnnotationDbi_1.42.1
##  [7] IRanges_2.14.12      S4Vectors_0.18.3     graph_1.58.2        
## [10] viridis_0.5.1        viridisLite_0.3.0    KEGGprofile_1.22.0  
## [13] RCurl_1.95-4.12      bitops_1.0-6         NMF_0.21.0          
## [16] Biobase_2.40.0       BiocGenerics_0.26.0  cluster_2.0.9       
## [19] rngtools_1.3.1.1     pkgmaker_0.27        registry_0.5-1      
## [22] ggplot2_3.1.1        biomaRt_2.36.1       BiocStyle_2.8.2     
## [25] kableExtra_1.1.0     knitr_1.22           limma_3.36.5        
## [28] usethis_1.5.0        devtools_2.0.2       rmarkdown_1.12      
## 
## loaded via a namespace (and not attached):
##  [1] colorspace_1.4-1        rprojroot_1.3-2        
##  [3] XVector_0.20.0          fs_1.3.1               
##  [5] rstudioapi_0.10         remotes_2.0.4          
##  [7] bit64_0.9-7             ClusterR_1.1.9         
##  [9] KEGG.db_3.2.3           xml2_1.2.0             
## [11] codetools_0.2-16        doParallel_1.0.14      
## [13] pkgload_1.0.2           ade4_1.7-13            
## [15] gridBase_0.4-7          png_0.1-7              
## [17] FD_1.0-12               readr_1.3.1            
## [19] compiler_3.6.0          httr_1.4.0             
## [21] backports_1.1.4         Matrix_1.2-17          
## [23] assertthat_0.2.1        lazyeval_0.2.2         
## [25] cli_1.1.0               htmltools_0.3.6        
## [27] prettyunits_1.0.2       tools_3.6.0            
## [29] gmp_0.5-13.5            gtable_0.3.0           
## [31] glue_1.3.1              reshape2_1.4.3         
## [33] dplyr_0.8.0.1           BiocWorkflowTools_1.6.2
## [35] Rcpp_1.0.1              Biostrings_2.48.0      
## [37] NMI_2.0                 ape_5.3                
## [39] nlme_3.1-139            iterators_1.0.10       
## [41] xfun_0.6                stringr_1.4.0          
## [43] ps_1.3.0                testthat_2.1.1         
## [45] rvest_0.3.3             gtools_3.8.1           
## [47] XML_3.98-1.19           zlibbioc_1.26.0        
## [49] MASS_7.3-51.4           scales_1.0.0           
## [51] hms_0.4.2               curl_3.3               
## [53] yaml_2.2.0              memoise_1.1.0          
## [55] gridExtra_2.3           TeachingDemos_2.10     
## [57] stringi_1.4.3           RSQLite_2.1.1          
## [59] desc_1.2.0              foreach_1.4.4          
## [61] permute_0.9-5           pkgbuild_1.0.3         
## [63] bibtex_0.4.2            geometry_0.4.1         
## [65] rlang_0.4.0             pkgconfig_2.0.2        
## [67] matrixStats_0.54.0      evaluate_0.13          
## [69] lattice_0.20-38         purrr_0.3.2            
## [71] bit_1.1-14              processx_3.3.1         
## [73] tidyselect_0.2.5        plyr_1.8.4             
## [75] magrittr_1.5            bookdown_0.9           
## [77] R6_2.4.0                DBI_1.0.0              
## [79] mgcv_1.8-28             pillar_1.3.1           
## [81] withr_2.1.2             abind_1.4-5            
## [83] KEGGREST_1.20.2         tibble_2.1.1           
## [85] crayon_1.3.4            progress_1.2.0         
## [87] grid_3.6.0              blob_1.1.1             
## [89] callr_3.2.0             git2r_0.25.2           
## [91] vegan_2.5-4             digest_0.6.18          
## [93] webshot_0.5.1           xtable_1.8-4           
## [95] munsell_0.5.0           magic_1.5-9            
## [97] sessioninfo_1.1.1
\end{verbatim}

>>>>

@@

Time course data can either be on a single condition (to identify genes
changing over time) or or on multiple conditions (such as the influenza
dataset we are considering), which will alter slightly the types of questions
we are interested in.

@@

\begin{tabular}{l}
\toprule
Contrasts\\
\midrule
M.0-VL.0\\
M.3-VL.3\\
M.6-VL.6\\
M.9-VL.9\\
M.12-VL.12\\
\addlinespace
M.18-VL.18\\
M.24-VL.24\\
M.30-VL.30\\
M.36-VL.36\\
M.48-VL.48\\
\addlinespace
M.60-VL.60\\
M.72-VL.72\\
M.120-VL.120\\
M.168-VL.168\\
\bottomrule
\end{tabular}

@@

\begin{tabular}{lrrrrrr}
\toprule
  & M.0-VL.0\_pval & M.0-VL.0\_qval & M.0-VL.0\_lfc & M.3-VL.3\_pval & M.3-VL.3\_qval & M.3-VL.3\_lfc\\
\midrule
NM\_009912 & 0.273 & 0.475 & 2.202 & 0.767 & 0.876 & 1.158\\
NM\_008725 & 0.851 & 0.924 & 2.028 & 0.139 & 0.302 & 1.457\\
NM\_007473 & 0.088 & 0.219 & 0.678 & 0.917 & 0.959 & 0.764\\
ENSMUST00000094955 & 0.008 & 0.037 & 1.186 & 0.935 & 0.969 & -0.633\\
NM\_001042489 & 0.419 & 0.621 & 1.212 & 0.761 & 0.872 & 0.599\\
\addlinespace
NM\_008159 & 0.123 & 0.278 & 1.102 & 0.117 & 0.268 & 0.737\\
NM\_001013813 & 0.057 & 0.159 & 0.630 & 0.171 & 0.348 & 0.975\\
AK039774 & 0.015 & 0.059 & 0.637 & 0.062 & 0.171 & 0.971\\
NM\_013782 & 0.875 & 0.938 & 0.594 & 0.000 & 0.001 & 1.180\\
NM\_028622 & 0.767 & 0.876 & 1.648 & 0.946 & 0.974 & -1.865\\
\bottomrule
\end{tabular}

@@

\begin{tabular}{lrrrrr}
\toprule
  & M-K:1 & M-K:7 & M-K:11 & M-K:14 & M-K:2\\
\midrule
NM\_009912 & -0.0061424 & 0.0224948 & -0.0515947 & -0.1143973 & 0.0614619\\
NM\_008725 & 2.6547855 & 0.2734387 & -0.6882925 & 0.4718807 & -1.7463319\\
NM\_007473 & -0.1603623 & 0.1017655 & 0.5079573 & 0.0595773 & 0.4250185\\
ENSMUST00000094955 & 0.3505920 & 0.6463011 & 0.2564796 & 0.3599102 & 0.4177116\\
NM\_001042489 & 0.0889750 & 0.2321381 & -0.1891501 & 0.1786113 & 0.3320644\\
\addlinespace
NM\_008159 & -0.3863770 & 0.0298395 & -0.0862755 & -0.3266351 & -0.1893062\\
\bottomrule
\end{tabular}

@@

\begin{tabular}{llrr}
\toprule
  & Pathway & Percentage & Adj. p-value\\
\midrule
04060 & Cytokine-cytokine receptor interaction & 34 & 0\\
04620 & Toll-like receptor signaling pathway & 40 & 0\\
04630 & Jak-STAT signaling pathway & 33 & 0\\
03050 & Proteasome & 53 & 0\\
04621 & NOD-like receptor signaling pathway & 47 & 0\\
\addlinespace
04380 & Osteoclast differentiation & 34 & 0\\
04623 & Cytosolic DNA-sensing pathway & 45 & 0\\
04210 & Apoptosis & 37 & 0\\
04622 & RIG-I-like receptor signaling pathway & 39 & 0\\
05160 & Hepatitis C & 29 & 0\\
\bottomrule
\end{tabular}

@@

\begin{tabular}{lllrrrlr}
\toprule
  & GO ID & Description & Annotated & Significant & Expected & P-value & adj. p-value\\
\midrule
5 & GO:0003002 & regionalization & 172 & 153 & 134.41 & 0.00026 & 0.2629\\
6 & GO:0048562 & embryonic organ morphogenesis & 152 & 135 & 118.78 & 0.00045 & 0.3323\\
7 & GO:0010970 & transport along microtubule & 82 & 75 & 64.08 & 0.00046 & 0.3323\\
8 & GO:0006805 & xenobiotic metabolic process & 49 & 47 & 38.29 & 0.00059 & 0.3729\\
9 & GO:0060173 & limb development & 92 & 84 & 71.89 & 0.00068 & 0.3820\\
\addlinespace
10 & GO:0019395 & fatty acid oxidation & 53 & 51 & 41.42 & 0.00134 & 0.6194\\
11 & GO:0055114 & oxidation-reduction process & 536 & 461 & 418.84 & 0.00139 & 0.6194\\
12 & GO:0060070 & canonical Wnt signaling pathway & 149 & 131 & 116.43 & 0.00147 & 0.6194\\
13 & GO:0090596 & sensory organ morphogenesis & 134 & 118 & 104.71 & 0.00219 & 0.8517\\
14 & GO:0060411 & cardiac septum morphogenesis & 33 & 32 & 25.79 & 0.00295 & 1.0000\\
\addlinespace
15 & GO:0010811 & positive regulation of cell-substrate ad... & 81 & 73 & 63.30 & 0.00384 & 1.0000\\
16 & GO:0042738 & exogenous drug catabolic process & 22 & 22 & 17.19 & 0.00437 & 1.0000\\
17 & GO:1901381 & positive regulation of potassium ion tra... & 22 & 22 & 17.19 & 0.00437 & 1.0000\\
18 & GO:0007368 & determination of left/right symmetry & 66 & 60 & 51.57 & 0.00536 & 1.0000\\
19 & GO:0060078 & regulation of postsynaptic membrane pote... & 66 & 60 & 51.57 & 0.00536 & 1.0000\\
\bottomrule
\end{tabular}

@@

\begin{verbatim}
## R version 3.6.0 (2019-04-26)
## Platform: x86_64-pc-linux-gnu (64-bit)
## Running under: Ubuntu 19.04
## 
## Matrix products: default
## BLAS:   /usr/lib/x86_64-linux-gnu/blas/libblas.so.3.8.0
## LAPACK: /usr/lib/x86_64-linux-gnu/lapack/liblapack.so.3.8.0
## 
## locale:
## [1] C
## 
## attached base packages:
##  [1] stats4    parallel  splines   stats     graphics  grDevices utils    
##  [8] datasets  methods   base     
## 
## other attached packages:
##  [1] RColorBrewer_1.1-2   moanin_0.0.0         topGO_2.32.0        
##  [4] SparseM_1.77         GO.db_3.6.0          AnnotationDbi_1.42.1
##  [7] IRanges_2.14.12      S4Vectors_0.18.3     graph_1.58.2        
## [10] viridis_0.5.1        viridisLite_0.3.0    KEGGprofile_1.22.0  
## [13] RCurl_1.95-4.12      bitops_1.0-6         NMF_0.21.0          
## [16] Biobase_2.40.0       BiocGenerics_0.26.0  cluster_2.0.9       
## [19] rngtools_1.3.1.1     pkgmaker_0.27        registry_0.5-1      
## [22] ggplot2_3.1.1        kableExtra_1.1.0     biomaRt_2.36.1      
## [25] BiocStyle_2.8.2      knitr_1.22           limma_3.36.5        
## [28] usethis_1.5.0        devtools_2.0.2       rmarkdown_1.12      
## 
## loaded via a namespace (and not attached):
##  [1] colorspace_1.4-1        rprojroot_1.3-2        
##  [3] XVector_0.20.0          fs_1.3.1               
##  [5] rstudioapi_0.10         remotes_2.0.4          
##  [7] bit64_0.9-7             ClusterR_1.1.9         
##  [9] KEGG.db_3.2.3           xml2_1.2.0             
## [11] codetools_0.2-16        doParallel_1.0.14      
## [13] pkgload_1.0.2           ade4_1.7-13            
## [15] gridBase_0.4-7          png_0.1-7              
## [17] FD_1.0-12               readr_1.3.1            
## [19] compiler_3.6.0          httr_1.4.0             
## [21] backports_1.1.4         Matrix_1.2-17          
## [23] assertthat_0.2.1        lazyeval_0.2.2         
## [25] cli_1.1.0               htmltools_0.3.6        
## [27] prettyunits_1.0.2       tools_3.6.0            
## [29] gmp_0.5-13.5            gtable_0.3.0           
## [31] glue_1.3.1              reshape2_1.4.3         
## [33] dplyr_0.8.0.1           BiocWorkflowTools_1.6.2
## [35] Rcpp_1.0.1              Biostrings_2.48.0      
## [37] NMI_2.0                 ape_5.3                
## [39] nlme_3.1-139            iterators_1.0.10       
## [41] xfun_0.6                stringr_1.4.0          
## [43] ps_1.3.0                testthat_2.1.1         
## [45] rvest_0.3.3             gtools_3.8.1           
## [47] XML_3.98-1.19           zlibbioc_1.26.0        
## [49] MASS_7.3-51.4           scales_1.0.0           
## [51] hms_0.4.2               curl_3.3               
## [53] yaml_2.2.0              memoise_1.1.0          
## [55] gridExtra_2.3           TeachingDemos_2.10     
## [57] stringi_1.4.3           RSQLite_2.1.1          
## [59] desc_1.2.0              foreach_1.4.4          
## [61] permute_0.9-5           pkgbuild_1.0.3         
## [63] bibtex_0.4.2            geometry_0.4.1         
## [65] rlang_0.4.0             pkgconfig_2.0.2        
## [67] matrixStats_0.54.0      evaluate_0.13          
## [69] lattice_0.20-38         purrr_0.3.2            
## [71] bit_1.1-14              processx_3.3.1         
## [73] tidyselect_0.2.5        plyr_1.8.4             
## [75] magrittr_1.5            bookdown_0.9           
## [77] R6_2.4.0                DBI_1.0.0              
## [79] pillar_1.3.1            withr_2.1.2            
## [81] mgcv_1.8-28             abind_1.4-5            
## [83] KEGGREST_1.20.2         tibble_2.1.1           
## [85] crayon_1.3.4            progress_1.2.0         
## [87] grid_3.6.0              blob_1.1.1             
## [89] callr_3.2.0             git2r_0.25.2           
## [91] vegan_2.5-4             digest_0.6.18          
## [93] webshot_0.5.1           xtable_1.8-4           
## [95] munsell_0.5.0           magic_1.5-9            
## [97] sessioninfo_1.1.1
\end{verbatim}
